Global land cover fraction mapping: methods, variables, challenges

### 2020-04-29 article/article.tex
Make abstract more concise

\abstract{Currently most global land cover maps represent land cover classes as discrete units. However, in reality pixels often contain mixed land cover. This causes loss of precision as only the dominant land cover class may be represented. A common workaround to this issue is to define mosaic classes that are combinations of pure land cover classes mixed with certain ratios, e.g. open forest as a combination of grassland and trees. However, this approach leads to increasingly complex map legends and confusion about the definitions.

Our proposed solution is to map land cover fractions: define the proportion that each land cover class occupies within the pixel. Such fuzzy classification both increases precision of the map, and also gives users the flexibility to derive their own maps with a customised legend, rather than relying on the definition that the mapmakers provide. Thus exhaustive land cover fraction maps are desired by a number of user communities.

We compared the performance of eight machine learning regression algorithms for land cover fraction mapping. We used and assessed six types of covariates for training the models: spectral bands and vegetation indices derived from Proba-V 100 m TOC reflectance; temporal features from its time series; terrestrial features from ASTER GDEM; soil properties from SoilGrids; climate data from WorldClim; and location. The models were trained on 138164 training points throughout the globe, describing land cover fractions of pixels aligned to the Proba-V 100 m grid. After training, their land cover class fraction predictions were compared with an independent validation set of 20705 points, likewise aligned to the Proba-V grid.

One challenge for fraction mapping models is data sparsity: most fractions are 0\%, but regression favours the mean, thus 0\% and 100\% are difficult to predict well. We propose a new solution by combining three models. A binary model determines whether a pixel is pure; if so, it is processed using a classification model; otherwise with a regression model.

Results show that the three-model approach results in better estimation of extreme values (lower MAE of 8.4 vs 9.4) at the cost of misclassification (higher RMSE of 19.4 vs 17.3). From the compared models, random forest and cubist regressions reach the highest overall accuracy of 72\%, comparable to the reported accuracies of the current discrete global land cover maps. The models were driven primarily by spectral remote sensing data, followed by soil and climate data.

This research proves that machine learning algorithms can be applied globally to a wide variety of land cover to create an exhaustive global land cover fraction map. This expresses land cover more precisely, and empowers users to create their own discrete maps, which helps land cover monitoring and climate modelling communities. In addition, this work contributes to efforts to produce exhaustive global land cover fractions as an operational product.

>>>>

\abstract{Currently most global land cover maps are produced with discrete classes, which express the dominant land cover class in each pixel, or a combination of several classes at a particular ratio. In contrast, land cover fraction mapping allows for expressing the proportion of each pure class in each pixel, which increases precision, reduces legend complexity, and allows users the flexibility of defining their own mosaic discrete classes using thresholds they define themselves. To map land cover fractions, regression rather than classification algorithms are needed, and multiple approaches are available for this task.

We compared the performance of eight machine learning regression algorithms for land cover fraction mapping, for the best model assessed the importance of six types of training variables, evaluated the spatial and per-class accuracy, and demonstrated a wall-to-wall prediction of seven land cover fractions over the globe.

The models were trained on over 130\,000 points and validated on a separate dataset of over 20\,000 points of the Copernicus Global Land System: Dynamic Land Cover (CGLS-LC100) project. Both datasets are global and aligned with the Proba-V 100 m grid.

One challenge for fraction mapping models is data sparsity: most fractions are 0\%, but regression favours the mean, thus 0\% and 100\% are difficult to predict well. We proposed a new solution by combining three models. A binary model determines whether a pixel is pure; if so, it is processed using a classification model; otherwise with a regression model.

Results showed that the three-model approach results in better estimation of extreme values (lower MAE of 8.4 vs 9.4) at the cost of misclassification (higher RMSE of 19.4 vs 17.3). From the compared models, random forest and cubist regressions reach the highest overall accuracy of 72\%, comparable to the reported accuracies of the current discrete global land cover maps. The models were driven primarily by spectral remote sensing data, followed by soil and climate data.

This research proves that machine learning algorithms can be applied globally to a wide variety of land cover to create an exhaustive global land cover fraction map. This expresses land cover more precisely, and empowers users to create their own discrete maps, which helps land cover monitoring and climate modelling communities. In addition, this work contributes to efforts to produce exhaustive global land cover fractions as an operational product, and to assess the accuracy of land cover fraction maps both thematically and spatially.
